---
title: "Hw02 -Shot Charts"
author: "Huiqing Fu"
date: "3/1/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../images/")
library(dplyr)
library(ggplot2)
library(magrittr)
library(jpeg)
library(grid)
```

## 1) File Structure (10 pts)
## 2) Data (10 pts)
####2.1) Download the data
```{r}
# curl -O https://raw.githubusercontent.com/ucb-stat133/stat133-spring-2018/master/data/stephen-curry.csv
# downloadgit     https://minhaskamal.github.io/DownGit/#/home
```
####2.2) Data Dictionary (10 pts)
## 3) Data Preparation (20 pts)
```{r}
setwd("../data")
```


```{r}
iguodala <- read.csv("../data/andre-iguodala.csv")
green<- read.csv("../data/draymond-green.csv")
curry<- read.csv("../data/stephen-curry.csv")
thompson<- read.csv("../data/klay-thompson.csv",stringsAsFactors = FALSE)
durant <- read.csv("../data/kevin-durant.csv",stringsAsFactors = FALSE)

# Add a column name to each imported data frame, that contains the name of the corresponding player
iguodala <- cbind(player="Andre Iguodala", iguodala )
green <- cbind(player="Draymond Green", green)
curry <- cbind(player="Stephen Curry", curry)
thompson <- cbind(player="Klay Thompson", thompson )
durant <- cbind(player="Kevin Durant", durant )
```
```{r}
#Change the original values of shot_made_flag to more descriptive values: replace "n" with "missed shot", and "y" with "made shot". Hint: you can use logical subsetting for this operation (no need to use programming structures that we haven¡¦t covered yet).
iguodala$shot_made_flag <- as.character(iguodala$shot_made_flag)
iguodala$shot_made_flag[iguodala$shot_made_flag == "n"] <- "missed shot"
iguodala$shot_made_flag[iguodala$shot_made_flag == "y"] <- "made shot"

green$shot_made_flag <- as.character(green$shot_made_flag)
green$shot_made_flag[green$shot_made_flag == "n"] <- "missed shot"
green$shot_made_flag[green$shot_made_flag == "y"] <- "made shot"

curry$shot_made_flag <- as.character(curry$shot_made_flag)
curry$shot_made_flag[curry$shot_made_flag == "n"] <- "missed shot"
curry$shot_made_flag[curry$shot_made_flag == "y"] <- "made shot"

thompson$shot_made_flag <- as.character(thompson$shot_made_flag)
thompson$shot_made_flag[thompson$shot_made_flag == "n"] <- "missed shot"
thompson$shot_made_flag[thompson$shot_made_flag == "y"] <- "made shot"

durant$shot_made_flag <- as.character(durant$shot_made_flag)
durant$shot_made_flag[durant$shot_made_flag == "n"] <- "missed shot"
durant$shot_made_flag[durant$shot_made_flag == "y"] <- "made shot"
```
```{r}
#Add a column minute that contains the minute number where a shot occurred. For instance, if a shot took place during period = 1 and minutes_remaining = 8, then this should correspond to a value minute = 4. Likewise, if a shot took place during period = 4 and minutes_remaining = 2 then this should correspond to a value minute = 46. Hint: you can use logical subsetting for these operations (no need to use programming structures that we haven¡¦t covered yet).
iguodala <- mutate(iguodala, minute=12*period-minutes_remaining )
green <- mutate(green, minute=12*period-minutes_remaining )
curry <- mutate(curry, minute=12*period-minutes_remaining )
thompson <- mutate(thompson, minute=12*period-minutes_remaining )
durant <- mutate(durant, minute=12*period-minutes_remaining )


```

```{r}
# Use sink() to send the summary() output of each imported data frame into individuals text files: andre-iguodala-summary.txt, draymond-green-summary.txt, etc. During each sinking operation, the produced summaries should be sent to the output/folder using relative paths.

sink("../output/andre-iguodala-summary.txt")
summary(iguodala)
sink()

sink("../output/klay-thompson-summary.txt")
summary(thompson)
sink()

sink("../output/draymond-green-summary.txt")
summary(green)
sink()

sink("../output/stephen-curry-summary.txt")
summary(curry)
sink()

sink("../output/kevin-durant-summary.txt")
summary(durant)
sink()
```

```{r}
# Use the row binding function rbind() to stack the tables into one single data frame (or tibble object).
dat <- rbind(iguodala, thompson, green, curry, durant)

# Export (i.e. write) the assembled table as a CSV file shots-data.csv inside the folder data/. Use a relative path for this operation.
write.csv(dat, file = "../data/shots-data.csv")

# Use sink() to send the summary() output of the assembled table. Send this output to a text file named shots-data-summary.txt inside the output/ folder. Use a relative path when exporting the R output.
sink("../output/shots-data-summary.txt")
summary(dat)
sink()
```

## 4) Shot Charts (20 pts)
####4.1) Shot charts of each player (10 pts)
```{r
# scatterplot
thompson_scatterplot <- ggplot(data = thompson) +
    geom_point(aes(x = x, y = y, color = shot_made_flag))
thompson_scatterplot
```
```{r}
#The JPG file with this image is available in the course github repository, inside the folder images/. You need to import the file nba-court.jpg to the images/ folder of your HW directory.
# cd ../images
#curl -O https://raw.githubusercontent.com/ucb-stat133/stat133-spring-2018/master/data/nba-court.jpg

# court image (to be used as background of plot)
court_file <- "../images/nba-court.jpg"
# create raste object

court_image <- rasterGrob(
  readJPEG(court_file),
  width = unit(1, "npc"),
  height = unit(1, "npc"))

# shot chart with court background
thompson_shot_chart <- ggplot(data = thompson) +
    annotation_custom(court_image, -250, 250, -50, 420) +
    geom_point(aes(x = x, y = y, color = shot_made_flag)) +
    ylim(-50, 420) +
    ggtitle('Shot Chart: Klay Thompson (2016 season)') +
    theme_minimal()
thompson_shot_chart 
```
####4.1) Shot charts of each player (10 pts)
Create shot charts (with court backgrounds) for each player, and save the plots in PDF
format, with dimensions width = 6.5 and height = 5 inches, inside the folder images/:
¡E andre-iguodala-shot-chart.pdf
¡E draymond-green-shot-chart.pdf
¡E kevin-durant-shot-chart.pdf
¡E klay-thompson-shot-chart.pdf
¡E stephen-curry-shot-chart.pdf
```{r iguodala_shot_chart}

#court_file <- "../images/nba-court.jpg"
court_image <- rasterGrob(readJPEG("../images/nba-court.jpg"),
                          width = unit(1,"npc"),
                          height = unit(1,"npc")
                    
)
iguodala_shot_chart <- ggplot(data = iguodala, aes(x = x, y =y))+
  annotation_custom(court_image, -250,250, -50, 420)+
  geom_point(aes(color=shot_made_flag))+
  ylim(-50,420)+
  ggtitle('Shot Chart: Andre Iguodala (2016 Season)')+
  theme_minimal()
 iguodala_shot_chart 
 
```
```{r green_shot_chart}
court_file <- "../images/nba-court.jpg"
court_image <- rasterGrob(
  readJPEG(court_file),
  width = unit(1, "npc"),
  height = unit(1, "npc")) 
green_shot_chart <- ggplot(data = green, aes(x = x, y = y, color = shot_made_flag))+
  annotation_custom(court_image,  -250, 250, -50,420)+
  ylim(-50,420)+
  geom_point()+
  ggtitle('Shot Chart: Draymond Green (2016 Season)')+
  theme_minimal()
green_shot_chart
```

```{r curry_shot_chart}
curry_shot_chart <- ggplot(data = curry, aes(x = x, y = y, color = shot_made_flag))+
  annotation_custom(court_image, -250,250,-50,420)+
  geom_point()+
  ylim(-50,420)+
  ggtitle("Shot Chart: Stephen Curry (2016 Season)")+
  theme_minimal()
curry_shot_chart
```
```{r durant_shot_chart}
durant_shot_chart <- ggplot(data = durant, aes(x = x, y = y, color = shot_made_flag))+
  annotation_custom(court_image, -250, 250, -50, 420)+
  geom_point()+
  ylim(-50,420)+
  ggtitle("Shot Chart: Kevin Durant (2016 Season)")+
  theme_minimal()
durant_shot_chart
```
```{r}
thompson_shot_chart <- ggplot(data = thompson, aes(x = x, y = y, color = shot_made_flag))+
  annotation_custom(court_image, -250, 250, -50, 420)+
  geom_point()+
  ylim(-50,420)+
  ggtitle("Shot Chart: Klay Thompson (2016 Season)")+
  theme_minimal()
thompson_shot_chart
```
```{r}
pdf(file = "../images/andre-iguodala-shot-chart.pdf", width = 7, height = 5)
iguodala_shot_chart
dev.off()

pdf(file = "../images/draymond-green-shot-chart.pdf", width = 7, height = 5)
green_shot_chart
dev.off()

pdf(file = "../images/kevin-durant-shot-chart.pdf", width = 7, height = 5)
durant_shot_chart
dev.off()

pdf(file = "../images/klay-thompson-shot-chart.pdf", width = 7, height = 5)
thompson_shot_chart
dev.off()

pdf(file = "../images/stephen-curry-shot-chart.pdf", width = 7, height = 5)
curry_shot_chart
dev.off()
```

####4.2) Facetted Shot Chart (10 pts)
Create one graph, using facetting, to show all the shot charts in one image, similar to the one
below. Save this image in PDF format as gsw-shot-charts.pdf, inside the folder images/.
Specify image dimensions width = 8 and height = 7 inches.
```{r}
gsw_shot_charts<- ggplot(data = dat, aes(x = x, y =y, color = shot_made_flag))+
         annotation_custom(court_image, -250, 250, -50, 420)+
  geom_point(size = 0.7)+
  ylim(-50, 420)+
  facet_wrap(~player)+
  ggtitle("Shot Charts: GSW (2016 season)")+
  theme_minimal()+
  theme(legend.position="top")+
  theme(legend.title = element_blank())+
  #theme(legend.position = c(0.8, 0.3))+
   theme(legend.text = element_text(size=8))
gsw_shot_charts
```
```{r}
pdf("../images/gsw-shot-charts.pdf", width = 8, height = 7)
gsw_shot_charts

dev.off()
```
## 5) Summary Tables (20 pts)
####5.1) Total Shots by Player (10 pts)
```{r}
# Total number of shots (2PT and 3PT, both made and missed) by player, arranged in descending order.
dat_total <- dat %>% 
  group_by(player)%>%
  summarise(count=n())%>%
  arrange(desc(count))
colnames(dat_total)[colnames(dat_total)=="count"] <- "total"
dat_total
```


####5.2) Effective Shooting Percentage (10 pts)
```{r}
# Effective Shooting % by Player, arranged in descending order.
dat_total <- dat %>%
  group_by(player)%>%
  summarise(count=n())
colnames(dat_total)[colnames(dat_total)=="count"] <- "total"
 
effective_shot <- dat%>%
  group_by(player,shot_made_flag)%>%
  summarise(count=n())%>%
  filter(shot_made_flag=="made shot")
colnames(effective_shot)[colnames(effective_shot)=="count"] <- "made"

EF_ST <- merge(dat_total, effective_shot, by = "player") %>%
  mutate(perc_made = made/total)%>%
  select(player, total, made,perc_made)%>%
  arrange(desc(perc_made))
EF_ST
```

```{r}
#2PT Effective Shooting % by Player:arranged in descending order by percentage.
dat_total_2PT <- dat %>%
  select(player, shot_made_flag, shot_type) %>%
  filter(shot_type =="2PT Field Goal") %>%
  group_by(player) %>%
  summarise(count=n())
colnames(dat_total_2PT)[colnames(dat_total_2PT)=="count"] <- "total"

effective_shot_2PT <- dat %>%
  select(player, shot_made_flag, shot_type) %>%
  filter(shot_type=="2PT Field Goal"&shot_made_flag=="made shot")%>%
  group_by(player,shot_made_flag, shot_type)%>%
    summarise(count=n())
  colnames(effective_shot_2PT)[colnames(effective_shot_2PT)=="count"] <- "made"

EF_ST_2PT <- 
  merge(dat_total_2PT, effective_shot_2PT, by = "player")%>%
    mutate(perc_made_2PT = made/total)%>%
    arrange(desc(perc_made_2PT))
EF_ST_2PT
```

```{r}
#3PT Effective Shooting % by Player:arranged in descending order by percentage.
dat_total_3PT <- dat %>%
  select(player, shot_made_flag, shot_type) %>%
  filter(shot_type =="3PT Field Goal") %>%
  group_by(player) %>%
  summarise(count=n())
colnames(dat_total_3PT)[colnames(dat_total_3PT)=="count"] <- "total"

effective_shot_3PT <- 
  dat %>%
  select(player, shot_made_flag, shot_type) %>%
    filter(shot_made_flag=="made shot"&shot_type=="3PT Field Goal")%>%
    group_by(player,shot_type, shot_made_flag)%>%
    summarise(count=n())
  colnames(effective_shot_3PT)[colnames(effective_shot_3PT)=="count"] <- "made"

EF_ST_3PT <- 
  merge(dat_total_3PT, effective_shot_3PT, by = "player")%>%
    mutate(perc_made_3PT = made/total)%>%
    arrange(desc(perc_made_3PT))
EF_ST_3PT
```


## 6) Shooting Distance (20 pts)
####6.1) dplyr table (10 pts)
```{r}
#More precisely, use dplyr operations to obtain a tibble with two columns: shot_distance and made_shot_prop. The first row of the tibble should contain the value of distance = 0 ft, and the associated proportion of made shots (of all the five analyzed players). The second row should contain the value of distance = 1 ft, and the corresponding proportion of made shots; and so on.
Total_shot_distance <- dat%>%
  select(shot_distance,shot_made_flag)%>%
  group_by(shot_distance)%>%
  summarise(count=n())
colnames(Total_shot_distance)[colnames(Total_shot_distance)=="count"] <- "total"

made_shot <- 
  dat%>%
  select(shot_distance,shot_made_flag)%>%
  filter(shot_made_flag=="made shot")%>%
  group_by(shot_distance)%>%
  summarise(count=n())
  colnames(made_shot)[colnames(made_shot)=="count"] <- "made"
  
Shooting_Distance <- merge(Total_shot_distance, made_shot, by = "shot_distance")%>%
  mutate(made_shot_prop = made/total)%>%
  select(shot_distance, made_shot_prop)
Shooting_Distance
```

####6.2) ggplot (10 pts)
```{r}
#With the tibble created in 6.1), use ggplot() to make a scatterplot with the variables shot_distance and made_shot_prop. Use the x-axis for the shot distance, and the y-axis for the proportion of made shots.
ggplot(data = Shooting_Distance, aes(x = shot_distance, y = made_shot_prop))+
  geom_point(aes(color=shot_distance))

```


¡E What do you observe?
Between 0-5 feet,it seems there is a linear relationship between shot_distance and made_shot_prop. However, between 5-35 feet, the points are bouncing around, it is a very weak correlation. At distance = 50, there is a point which is clearly a outlier. Therefore, we have no a clear conclusion for NBA players that the shorter the shooting distance, the higher the chance to successfully make a shot. 

¡E Can you confirm that the shorter the distance, the more effective the shots?
Based on the data, we cannot have such a conclusion for NBA players. 

¡E Can you guesstimate a distance threshold beyond which the chance of making a successful shot is basically null?
Beyond 5 feet,  the chance of making a successful shot is basically null.

¡E What distances tend to have a percentage (of making a shot) of 50% or more?
15-22 feet.

## 7) Total number of shots by minute of occurrence (10 pts)

```{r}
#total number of shots (made and missed),minute of occurrence.
#¡E stephen-curry
shots_curry <- curry %>%
  select(player, minute)%>%
  group_by(player,minute)%>%
  summarise(count=n())%>%
  ggplot(aes(x = minute, y = count),color='blue' )+
  geom_point()+
  ylab("total number of shots")+
  scale_x_continuous(limits = c(0,48), breaks = seq(from=2, to =48, by = 2))+
  scale_y_continuous(limits = c(0,70))+
  geom_path()+
  geom_rect(alpha= 0.005,aes(x=minute, y =count,xmin = 0, xmax=12, ymin=-Inf, ymax=Inf))+
geom_rect(alpha=0.005,aes(x=minute, y =count,xmin = 24, xmax=36, ymin=-Inf, ymax=Inf))+
  theme_minimal()+
  ggtitle("Stephen Curry: Total Number of Shots")
shots_curry

#¡E andre-iguodala
shots_iguodala <- iguodala %>%
  select(player, minute)%>%
  group_by(player,minute)%>%
  summarise(count=n())%>%
  ggplot(aes(x = minute, y = count))+
  geom_point()+
  ylab("total number of shots")+
  scale_x_discrete(limits = c(0,48), breaks = seq(from=2, to =48, by = 2))+
  scale_y_discrete(limits = c(0,70))+
  geom_path()+
  geom_rect(alpha= 0.005,fill="blue",aes(x=minute, y =count,xmin = 0, xmax=12, ymin=-Inf, ymax=Inf))+
  geom_rect(alpha=0.005,fill="blue",aes(x=minute, y =count,xmin = 24, xmax=36, ymin=-Inf, ymax=Inf))+
  theme_minimal()+
  ggtitle("Andre Iguodala: Total Number of Shots")
shots_iguodala

#¡E draymond-green
shots_green <- green %>%
  select(player, minute)%>%
  group_by(player,minute)%>%
  summarise(count=n())%>%
  ggplot(aes(x = minute, y = count) )+
  geom_point(color='maroon')+
  ylab("total number of shots")+
  scale_x_discrete(limits = c(0,48), breaks = seq(from=2, to =48, by = 2))+
  scale_y_discrete(limits = c(0,70))+
  geom_path(color='coral1')+
  geom_rect(alpha= 0.005,fill="yellowgreen",aes(x=minute, y =count,xmin = 0, xmax=12, ymin=-Inf, ymax=Inf))+
geom_rect(alpha=0.005,fill="yellowgreen",aes(x=minute, y =count,xmin = 24, xmax=36, ymin=-Inf, ymax=Inf))+
  theme_minimal()+
  ggtitle("Draymond Green: Total Number of Shots")
shots_green

#¡E kevin-durant
shots_durant <- durant %>%
  select(player, minute)%>%
  group_by(player,minute)%>%
  summarise(count=n())%>%
  ggplot(aes(x = minute, y = count))+
  geom_point(color="deepskyblue3")+
  ylab("total number of shots")+
  scale_x_continuous(limits = c(0,48), breaks = seq(from=2, to =48, by = 2))+
  scale_y_continuous(limits = c(0,70))+
  geom_path(color="deepskyblue2")+
  geom_rect(alpha= 0.005,aes(x=minute, y =count),xmin = 0, xmax=12, ymin=-Inf, ymax=Inf)+
geom_rect(alpha=0.005,aes(x=minute, y =count),xmin = 24, xmax=36, ymin=-Inf, ymax=Inf)+
  theme_minimal()+
  ggtitle("Kevin Durant: Total Number of Shots")
shots_durant
#¡E klay-thompson
shots_thompson <- thompson %>%
  select(player, minute)%>%
  group_by(player,minute)%>%
  summarise(count=n())%>%
  ggplot(aes(x = minute, y = count),color='blue' )+
  geom_point(color="yellowgreen")+
  ylab("total number of shots")+
  scale_x_continuous(limits = c(0,48), breaks = seq(from=2, to =48, by = 2))+
  scale_y_continuous(limits = c(0,70))+
  geom_path(color="green")+
  geom_rect(alpha= 0.005,aes(x=minute, y =count),xmin = 0, xmax=12, ymin=-Inf, ymax=Inf)+
geom_rect(alpha=0.005,aes(x=minute, y =count),xmin = 24, xmax=36, ymin=-Inf, ymax=Inf)+
  theme_minimal()+
  ggtitle("Klay Thompson: Total Number of Shots")
shots_thompson
```
```{r}
#¡E Total Number of Shots (made and missed),minute of occurrence

 dat %>%
  select(player, minute)%>%
  group_by(player,minute)%>%
  summarise(count=n())%>%
  ggplot(aes(x = minute, y = count),color='blue')+
  geom_point()+
  ylab("total number of shots")+
  scale_x_continuous(limits = c(0,48), breaks = c(1,12,24,36,48))+
  scale_y_continuous(limits = c(0,70))+
  geom_path()+
  geom_rect(alpha= 0.008,aes(x=minute, y =count, xmin = 0, xmax=12, ymin=-Inf, ymax=Inf), fill = "black")+
geom_rect(alpha=0.008,aes(x=minute, y =count,xmin = 24, xmax=36, ymin=-Inf, ymax=Inf),  fill = "black")+  #col = "grey",fill = "black"
  ggtitle("Total Number of Shots")+
  theme_minimal()+
  facet_wrap(~player)
```








